---
title: "Untitled"
author: ""
date: "November 30, 2017"
output: pdf_document
---

#Introduction to Dengue

Introduction:
    Incidence of Dengue follows seasonal transmission patterns, and tend to have outbreaks every fews years. Since these epidemics are currently hard to predict and they affect to the population in areas where dengue is endemic. Early recognition and prompt treatment of severe cases can substantially lower the risk of medical complications and death. Accurate forecasts of cases of infected individuals, or incidence, are key to planning and resource allocation. For example, knowing well in advance the numbers of cases that are expected and when they will occur allows preparation via education and community mobilization campaigns, reallocation of resources (people, insecticide, diagnostic reagents) to high-risk areas, or re-training of physicians to recognize symptoms and to treat appropriately (Kuhn et al., 2005; Degallier et al., 2010; Thomson et al., 2008) in advance of peak transmission. Our goal is to develop an optimal dlm model to predict the epidemics of dengue using only data from time periods prior to the historical dengue seasons. 

Data:
    The provided data include weekly dengue incidence and linked environmental variables, and the training and testing sets may be downloaded from http://dengueforecasting.noaa.gov/ The dengue incidence portion is comprised of historical surveillance data at Iquitos, Peru and San Juan, Puerto Rico, summarized weekly. Cases in the data set include laboratory-confirmed and serotype-specific cases. The data are reflecting the total number of cases in each week, possibly revised or estimated ex post. A breakdown of incidence into strata of four serotypes, with a fifth un-serotyped category, were also provided. However, we only trained on total_cases in the data file.
#Basic attempts at modelling 

##Locally-level
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]
plot.ts(dat$total_cases, ylab="Incidence")

#LOCAL LEVEL BUILD
locally_level_build <- function(x){
  dlmModPoly(order=1,dV=exp(x[1]),dW=exp(x[2]))
}

#COMPUTE MLE ESTIMATE
locally_level_mle <- dlmMLE(dat$total_cases, parm=c(0,0), build=locally_level_build)

#GET FILTERED RESULTS
locally_level_filtered <- dlmFilter(dat$total_cases, locally_level_build(locally_level_mle$par))

#PLOT FILTERED RESULTS AND DIAGNOSTICS
plot(locally_level_filtered$f)
qqnorm(residuals(locally_level_filtered, sd=FALSE))
qqline(residuals(locally_level_filtered,sd=FALSE))
```

##Locally-level+Seasonal
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]
plot.ts(dat$total_cases, ylab="Incidence")

#LOCAL LEVEL BUILD
locally_level_seasonal_build <- function(x){
  dlmModPoly(order=1,dV=exp(x[1]),dW=exp(x[2])) +
    dlmModSeas(frequency = 4, dV = 0, dW = c(exp(x[3]),rep(0,2)))
}

#COMPUTE MLE ESTIMATE
locally_level_seasonal_mle <- dlmMLE(dat$total_cases, parm=c(0,0,0), build=locally_level_seasonal_build)

#GET FILTERED RESULTS
locally_level_seasonal_filtered <- dlmFilter(dat$total_cases, locally_level_seasonal_build(locally_level_seasonal_mle$par))


plot(locally_level_seasonal_filtered$f)
qqnorm(residuals(locally_level_seasonal_filtered, sd=FALSE))
qqline(residuals(locally_level_seasonal_filtered,sd=FALSE))
```

##Locally-linear + Seasonal
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]
plot.ts(dat$total_cases, ylab="Incidence")

#LOCAL LEVEL BUILD
locally_linear_seasonal_build <- function(x){
  dlmModPoly(order=2,dV=exp(x[1]),dW=exp(x[2])) +
    dlmModSeas(frequency = 52, dV = 0, dW = c(exp(x[3]),rep(0,50)))
}

#COMPUTE MLE ESTIMATE
locally_linear_seasonal_mle <- dlmMLE(dat$total_cases, parm=c(0,0,0), build=locally_level_seasonal_build)

#GET FILTERED RESULTS
locally_linear_seasonal_filtered <- dlmFilter(dat$total_cases, locally_level_seasonal_build(locally_linear_seasonal_mle$par))


plot(locally_linear_seasonal_filtered$f)
qqnorm(residuals(locally_linear_seasonal_filtered, sd=FALSE))
qqline(residuals(locally_linear_seasonal_filtered,sd=FALSE))

```

##Seasonal vs Fourier Transform
```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]
plot.ts(dat$total_cases, ylab="Incidence")

locally_level_fourier_build <- function ( x ) {
  dlmModPoly ( order=1, dV=exp(x[1]), dW=0 ) +
    dlmModTrig ( tau=52, q=1, dV=0, dW=rep(exp(x[2]),2) )
}

#COMPUTE MLE ESTIMATE
locally_level_fourier_mle <- dlmMLE(dat$total_cases, parm=c(0,0), build=locally_level_fourier_build)

#GET FILTERED RESULTS
locally_level_fourier_filtered <- dlmFilter(dat$total_cases, locally_level_fourier_build(locally_level_fourier_mle$par))

plot(locally_level_fourier_filtered$f)
qqnorm(residuals(locally_level_fourier_filtered, sd=FALSE))
qqline(residuals(locally_level_fourier_filtered,sd=FALSE))

```

##Multiple seasonality 
```{r, message=FALSE, warning=FALSE, echo=FALSE}

library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]
plot.ts(dat$total_cases, ylab="Incidence")

#LOCAL LEVEL BUILD
locally_linear_seasonal_build <- function(x){
  dlmModPoly(order=2,dV=exp(x[1]),dW=exp(x[2])) +
    dlmModSeas(frequency = 8, dV = 0, dW = c(exp(x[3]),rep(0,6))) +
       dlmModSeas(frequency = 4, dV = 0, dW = c(exp(x[4]),rep(0,2))) 

}

#COMPUTE MLE ESTIMATE
locally_linear_seasonal_mle <- dlmMLE(dat$total_cases, parm=c(0,0,0,0), build=locally_level_seasonal_build)

#GET FILTERED RESULTS
locally_linear_seasonal_filtered <- dlmFilter(dat$total_cases, locally_level_seasonal_build(locally_linear_seasonal_mle$par))


plot(locally_linear_seasonal_filtered$f)
qqnorm(residuals(locally_linear_seasonal_filtered, sd=FALSE))
qqline(residuals(locally_linear_seasonal_filtered,sd=FALSE))
```




#Alpha model

##Standard

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]
plot.ts(dat$total_cases, ylab="Incidence")
```
\bigskip

$$Y_t = F_t\Theta_t + W_t$$
$$\Theta_t = G\Theta_{t-1} + V_t$$

When t is not in summer, $F_t=F=1$,\par
when t is in summer, $F_t=\alpha_sF$.\par
where $\alpha_s$ is the maximum incidence in each year, $s=1,...,20$. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
a<-c()
t<-c()

for (i in 1:19) {
  a[i] <- max(dat[((i-1)*52+1):(i*52),2])
  t[i] <- which.max(dat[((i-1)*52+1):(i*52),2])
}

X <- c()
for (i in 1:19){
  X <-c(X,c(rep(1,22),rep(a[i],13),rep(1,17)))
}
#X <- rep(c(rep(1,22),rnorm(13,median(a),1),rep(1,17)),19)
X <- cbind(X,rep(0,988))

GG <- matrix(rep(0,16),nrow=4,ncol=4)
GG[1,1] = 1
GG[3,3] = cos(2*pi/52)
GG[3,4] = sin(2*pi/52)
GG[4,3] = -sin(2*pi/52)
GG[4,4] = cos(2*pi/52)

mod1 <- function(x){dlm(m0=rep(0,4), C0=1e+07*diag(4), 
                        FF=matrix(c(1,0,1,0),1,4),GG=GG, 
                        V=exp(x[1]), W=diag(c(exp(x[2]),exp(x[3]),exp(x[4]),exp(x[5]))), 
                        X=X, JFF=matrix(c(1,0,0,0),1,4))}

mle <- dlmMLE(dat$total_cases, parm = rep(1, 5), build=mod1)

filt = dlmFilter(y=dat$total_cases,mod =mod1(mle$par))
plot.ts(dat$total_cases, ylab="Incidence") # filtered
lines(dropFirst(filt$m[,1])+dropFirst(filt$m[,3]),col="red") # original data
legend("topright", legend=c("filtered","data"), col=c("red","black"), lty=1)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.ts(dat$total_cases, ylab="Incidence") # filtered
lines(dropFirst(filt$m[,1])+dropFirst(filt$m[,3]),col="red") # original data
legend("topright", legend=c("filtered","data"), col=c("red","black"), lty=1)
abline(v=75,col="grey");abline(v=87,col="grey");
abline(v=231,col="grey");abline(v=243,col="grey");
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.ts(dat$total_cases,ylab="incidence")
lines(filt$f,col="red")
```

##log-alpha model 
```{r}
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]

cases <- dat$total_cases[1:100]
require(rbiips)
library(MCMCpack)
model_file = '/Users/gcgibson/DLM_final_project/log_alpha.bug' # BUGS model filename
cat(readLines(model_file), sep = "\n")

par(bty='l')
light_blue = rgb(.7, .7, 1)
light_red = rgb(1, .7, .7)


t_max = length(cases)
n_burn = 500 # nb of burn-in/adaptation iterations
n_iter = 1000 # nb of iterations after burn-in
thin = 5 # thinning of MCMC outputs
n_part = 50 # nb of particles for the SMC
param_names = c('alpha') # names of the variables updated with MCMC (others are updated with SMC)
latent_names = c('x') # names of the variables updated with SMC and that need to be monitored

G = matrix(c(cos(2*pi),sin(2*pi),-sin(2*pi),cos(2*pi)), nrow=2, byrow=TRUE)

#setting the mean value of the initial count to 1400
data = list(t_max=t_max, y = cases,  G = G, mean_sigma_init = c(0,0), cov_sigma_init=diag(2) ,mean_x_init=c(0,0))
model = biips_model(model_file, data=data,sample_data = FALSE)

##fixing variance for now, will extend model to handle inference over variance later
n_burn = 5000 # nb of burn-in/adaptation iterations
n_iter = 10000 # nb of iterations after burn-in
thin = 5 # thinning of MCMC outputs
n_part = 50 # nb of particles for the SMC
param_names = c('alpha') # names of the variables updated with MCMC (others are updated with SMC)
latent_names = c('x') # names of the variables updated with SMC and that need to be monitored
inits = list(10)

obj_pmmh = biips_pmmh_init(model, param_names, inits=list(1),
                           latent_names=latent_names) # creates a pmmh object
biips_pmmh_update(obj_pmmh, n_burn, n_part) # adaptation and burn-in iterations
out_pmmh = biips_pmmh_samples(obj_pmmh, n_iter, n_part, thin=thin) # samples

summ_pmmh = biips_summary(out_pmmh, probs=c(.025, .975))

```

##Markov chain switching model
```{r}
model_file = '/Users/gcgibson/DLM_Final_Project/switching.bug' # BUGS model filename
cat(readLines(model_file), sep = "\n")

require(rbiips)
par(bty='l')
light_blue = rgb(.7, .7, 1)
light_red = rgb(1, .7, .7)
hot_colors = colorRampPalette(c('black', 'red', 'yellow', 'white'))

set.seed(0)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]

cases <- dat$total_cases[seq(1,10)]

sigma = .4; alpha = c(-2.5, -1); phi = .5; c0 = 1; x0 = 0; t_max = length(cases)
pi = matrix(c(.9, .1, .1, .9), nrow=2, byrow=TRUE)
data = list(t_max=t_max, sigma=sigma,y=cases,
            alpha=alpha, phi=phi, pi=pi, c0=c0, x0=x0)

model = biips_model(model_file, data, sample_data=FALSE)

n_burn = 2000
n_iter = 10000
thin = 1
n_part = 50


obj_pimh = biips_pimh_init(model, variables)


biips_pimh_update(obj_pimh, n_burn, n_part) # Burn-in iterations

out_pimh = biips_pimh_samples(obj_pimh, n_iter, n_part, thin=thin) # Return samples
summ_pimh = biips_summary(out_pimh, probs=c(.025, .975))
```



##Mechanistic Integration
$$\beta \sim Exp(\lambda_1)$$
$$\gamma \sim Exp(\lambda_2)$$
$$(S_t,I_t,R_t) \sim Dirichlet(f(S_{t-1},I_{t-1},R_{t-1},\beta,\gamma))$$



$$f = \text{rkf approx to ode}$$



$$Y_t \sim Pois(NI_t)$$








```{r, message=FALSE,warning=FALSE,echo=FALSE}
cases <- dat$total_cases[1:200]
require(rbiips)
library(MCMCpack)
dMN_dim <- function(s,i,r) {
  # Check dimensions of the input and return dimension of the output of
  # distribution dMN
  3
}
dMN_sample <- function(s,i,r) {
  # Draw a sample of distribution dMN

  rsamp <- rmultinom(20000,c(s/(s+i+r),i/(s+i+r),r/(s+i+r)))
  c(rsamp[1],rsamp[2],rsamp[3])
}
biips_add_distribution('ddirch', 3, dMN_dim, dMN_sample)


if (FALSE){
model_file = '/Users/gcgibson/dlm_final_project/blob.bug' # BUGS model filename
cat(readLines(model_file), sep = "\n")

par(bty='l')
light_blue = rgb(.7, .7, 1)
light_red = rgb(1, .7, .7)

t_max = length(cases)
mean_x_init = 1
prec_x_init = 1/5
prec_x = 1/10
log_prec_y_true = log(1) # True value used to sample the data
data = list(t_max=t_max, y = cases +1,prec_x_init=prec_x_init,
            prec_x=prec_x, 
            mean_x_init=c(.9,.05,.05))

sample_data = FALSE # Boolean
model = biips_model(model_file, data, sample_data=sample_data) # Create Biips model and sample data

data = model$data()

### PMMH

  n_burn = 2000 # nb of burn-in/adaptation iterations
  n_iter = 2000 # nb of iterations after burn-in
  thin = 1 # thinning of MCMC outputs
  n_part = 50 # nb of particles for the SMC
  param_names = c('beta','gamma') # name of the variables updated with MCMC (others are updated with SMC)
  latent_names = c('x') # name of the variables updated with SMC and that need to be monitored
  
  
  obj_pmmh = biips_pmmh_init(model, param_names, inits=list(beta=.9,gamma=.1),
                             latent_names=latent_names) # creates a pmmh object
  biips_pmmh_update(obj_pmmh, n_burn, n_part) # adaptation and burn-in iterations
  
  out_pmmh = biips_pmmh_samples(obj_pmmh, n_iter, n_part, thin=thin) # samples
  
  summ_pmmh = biips_summary(out_pmmh, probs=c(.025, .975))




### PARTICLE FILTER
n_part = 100000 # Number of particles
variables = c('x','y') # Variables to be monitored
mn_type = 'fs'; rs_type = 'stratified'; rs_thres = 0.9 # Optional parameters



out_smc = biips_smc_samples(model, variables, n_part,
                            type=mn_type, rs_type=rs_type, rs_thres=rs_thres)
diag_smc = biips_diagnosis(out_smc)
summ_smc = biips_summary(out_smc, probs=c(.025, .975))


x_f_mean = summ_smc$x$f$mean
x_f_quant = summ_smc$x$f$quant


filtered_states = 2000*x_f_mean[2,]
last_state = tail(t(x_f_mean),n=1)


forecasts <- c()
for (i in seq(1,52)){
  last_state[1] = last_state[1] +1/6*(-2*last_state[1]*last_state[2])
  last_state[2] =last_state[2]+ 1/6*(2*last_state[1]*last_state[2]-1.4*last_state[2])
  last_state[3] = last_state[3] + 1/6*(1.4*last_state[2])
  last_state <- rdirichlet(1,1e4*last_state)
  forecasts <- c(forecasts,last_state[2])
}

forecasts <- 2000*forecasts
plot(forecasts)
}


```
