---
title: "Untitled"
author: ""
date: "November 30, 2017"
output: pdf_document
---

#Introduction to Dengue


#Basic attempts at modelling 
##Locally-level
##Locally-level+Seasonal
##Locally-linear+ Seasonal
##Seasonal vs Fourier Transform
##Multiple seasonality 

#Alpha model
##Standard

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(dlm)
library(ggplot2)
SJdat<-read.csv("http://dengueforecasting.noaa.gov/Training/San_Juan_Training_Data.csv")
dat <- SJdat[c("season_week","total_cases")]

plot.ts(dat$total_cases, ylab="Incidence")
```
\bigskip

$$Y_t = F_t\Theta_t + W_t$$
$$\Theta_t = G\Theta_{t-1} + V_t$$

When t is not in summer, $F_t=F=1$,\par
when t is in summer, $F_t=\alpha_sF$.\par
where $\alpha_s$ is the maximum incidence in each year, $s=1,...,20$. 

```{r, message=FALSE, warning=FALSE, echo=FALSE}
a<-c()
t<-c()

for (i in 1:19) {
  a[i] <- max(dat[((i-1)*52+1):(i*52),2])
  t[i] <- which.max(dat[((i-1)*52+1):(i*52),2])
}

X <- c()
for (i in 1:19){
  X <-c(X,c(rep(1,22),rep(a[i],13),rep(1,17)))
}
#X <- rep(c(rep(1,22),rnorm(13,median(a),1),rep(1,17)),19)
X <- cbind(X,rep(0,988))

GG <- matrix(rep(0,16),nrow=4,ncol=4)
GG[1,1] = 1
GG[3,3] = cos(2*pi/52)
GG[3,4] = sin(2*pi/52)
GG[4,3] = -sin(2*pi/52)
GG[4,4] = cos(2*pi/52)

mod1 <- function(x){dlm(m0=rep(0,4), C0=1e+07*diag(4), 
                        FF=matrix(c(1,0,1,0),1,4),GG=GG, 
                        V=exp(x[1]), W=diag(c(exp(x[2]),exp(x[3]),exp(x[4]),exp(x[5]))), 
                        X=X, JFF=matrix(c(1,0,0,0),1,4))}

mle <- dlmMLE(dat$total_cases, parm = rep(1, 5), build=mod1)

filt = dlmFilter(y=dat$total_cases,mod =mod1(mle$par))
plot.ts(dat$total_cases, ylab="Incidence") # filtered
lines(dropFirst(filt$m[,1])+dropFirst(filt$m[,3]),col="red") # original data
legend("topright", legend=c("filtered","data"), col=c("red","black"), lty=1)
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.ts(dat$total_cases, ylab="Incidence") # filtered
lines(dropFirst(filt$m[,1])+dropFirst(filt$m[,3]),col="red") # original data
legend("topright", legend=c("filtered","data"), col=c("red","black"), lty=1)
abline(v=75,col="grey");abline(v=87,col="grey");
abline(v=231,col="grey");abline(v=243,col="grey");
```


```{r, message=FALSE, warning=FALSE, echo=FALSE}
plot.ts(dat$total_cases,ylab="incidence")
lines(filt$f,col="red")
```

##log-alpha model 


#Mechanistic Integration
$$\beta \sim Exp(\lambda_1)$$
$$\gamma \sim Exp(\lambda_2)
$$(S_t,I_t,R_t) \sim Dirichlet(f(S_{t-1},I_{t-1},R_{t-1},\beta,\gamma))$$



$$f = \text{rkf approx to ode}$$



$$Y_t \sim Pois(NI_t)$$








```{r, message=FALSE,warning=FALSE,echo=FALSE}
cases <- dat$total_cases[1:200]
require(rbiips)
library(MCMCpack)
dMN_dim <- function(s,i,r) {
  # Check dimensions of the input and return dimension of the output of
  # distribution dMN
  3
}
dMN_sample <- function(s,i,r) {
  # Draw a sample of distribution dMN

  rsamp <- rmultinomial(20000,c(s/(s+i+r),i/(s+i+r),r/(s+i+r)))
  c(rsamp[1],rsamp[2],rsamp[3])
}
biips_add_distribution('ddirch', 3, dMN_dim, dMN_sample)



model_file = '/Users/gcgibson/DLM_Final_Project/blob.bug' # BUGS model filename
cat(readLines(model_file), sep = "\n")

par(bty='l')
light_blue = rgb(.7, .7, 1)
light_red = rgb(1, .7, .7)

t_max = length(cases)
mean_x_init = 1
prec_x_init = 1/5
prec_x = 1/10
log_prec_y_true = log(1) # True value used to sample the data
data = list(t_max=t_max, y = cases +1,prec_x_init=prec_x_init,
            prec_x=prec_x, 
            mean_x_init=c(.9,.05,.05))

sample_data = FALSE # Boolean
model = biips_model(model_file, data, sample_data=sample_data) # Create Biips model and sample data

data = model$data()

### PMMH
if (FALSE){
  n_burn = 2000 # nb of burn-in/adaptation iterations
  n_iter = 2000 # nb of iterations after burn-in
  thin = 1 # thinning of MCMC outputs
  n_part = 50 # nb of particles for the SMC
  param_names = c('beta','gamma') # name of the variables updated with MCMC (others are updated with SMC)
  latent_names = c('x') # name of the variables updated with SMC and that need to be monitored
  
  
  obj_pmmh = biips_pmmh_init(model, param_names, inits=list(beta=.9,gamma=.1),
                             latent_names=latent_names) # creates a pmmh object
  biips_pmmh_update(obj_pmmh, n_burn, n_part) # adaptation and burn-in iterations
  
  out_pmmh = biips_pmmh_samples(obj_pmmh, n_iter, n_part, thin=thin) # samples
  
  summ_pmmh = biips_summary(out_pmmh, probs=c(.025, .975))
}



### PARTICLE FILTER
n_part = 10000 # Number of particles
variables = c('x','y') # Variables to be monitored
mn_type = 'fs'; rs_type = 'multinomial'; rs_thres = 0.9 # Optional parameters



out_smc = biips_smc_samples(model, variables, n_part,
                            type=mn_type, rs_type=rs_type, rs_thres=rs_thres)
diag_smc = biips_diagnosis(out_smc)
summ_smc = biips_summary(out_smc, probs=c(.025, .975))


x_f_mean = summ_smc$x$f$mean
x_f_quant = summ_smc$x$f$quant


filtered_states = 2000*x_f_mean[2,]
last_state = tail(t(x_f_mean),n=1)


forecasts <- c()
for (i in seq(1,52)){
  last_state[1] = last_state[1] +1/6*(-2*last_state[1]*last_state[2])
  last_state[2] =last_state[2]+ 1/6*(2*last_state[1]*last_state[2]-1.4*last_state[2])
  last_state[3] = last_state[3] + 1/6*(1.4*last_state[2])
  last_state <- rdirichlet(1,1e4*last_state)
  forecasts <- c(forecasts,last_state[2])
}

forecasts <- 2000*forecasts
plot(forecasts)



```
